---
title: "BloodTransfusionDataHarm"
author: "Yan Min"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Documents/CardioThoracicSurgeryResearch/BloodTransfusion")
```

# Load packages
```{r}
pkgs <- c("dplyr", "ggplot2", "tableone", "tidyr", "data.table")
sapply(pkgs, require, character.only = T)
```

# Extract the lowest postoperative hemoglobin
```{r}
cbc1 <- read.csv("AllCBCdata.csv", check.names = F) # cbc data for the 700 root surgeries
cbc1$surgdt <- cbc1$or_date
cbc1$or_date <- NULL
cbc2 <- read.csv("dalal_fiscbein_aortic_root_cohort_upd20210818.csv") # cbc data for all stanford sts cohort
cbc <- rbind(cbc1, cbc2) %>% as.data.frame()
remove(cbc1)
remove(cbc2)

cbc$ord_value <- cbc$ord_value %>% as.character() %>% as.numeric()
cbc_hgb <- cbc[cbc$lab_name == "Hemoglobin"&cbc$timing == "postop",]
# Seperate the blood taken date and time
cbc_hgb$taken_time <- as.character(cbc_hgb$taken_time)
cbc_hgb$taken_time <- as.POSIXct(cbc_hgb$taken_time, format="%m/%d/%y %H:%M")
cbc_hgb <- cbc_hgb[!duplicated(cbc_hgb$taken_time),]
mrns <- unique(cbc_hgb$mrn)

# Create the dataset of the lowest hgb level
cbc_lwst_hgb <- matrix(nrow = length(mrns), ncol = 2)
for(i in 1:length(mrns)){
  mrn <- mrns[i]
  sub <- cbc_hgb[cbc_hgb$mrn == mrn, ]
  lwst_hgb <- min(sub$ord_value)
  cbc_lwst_hgb[i,] <- c(mrn, lwst_hgb)
  remove(sub)
}
cbc_lwst_hgb <- as.data.frame(cbc_lwst_hgb)
colnames(cbc_lwst_hgb) <- c("mrn", "lwst_hgb")

#write.csv(cbc_lwst_hgb, file = "cbc_lwst_hgb_08202021.csv")

# Create the dataset of the first postoperative hgb level
cbc_fst_hgb <- matrix(nrow = length(mrns), ncol = 2)
for(i in 1:length(mrns)){
  mrn <- mrns[i]
  sub <- cbc_hgb[cbc_hgb$mrn == mrn, ]
  fst_hgb <- sub[sub$taken_time == min(sub$taken_time),]$ord_value
  cbc_fst_hgb[i,] <- c(mrn, fst_hgb)
  remove(sub)
}
cbc_fst_hgb <- as.data.frame(cbc_fst_hgb)
colnames(cbc_fst_hgb) <- c("mrn", "first_postophgb")

#write.csv(cbc_fst_hgb, file = "cbc_fst_postophgb_08202021.csv")
```

```{r}
mrn <- mrns[3]
  sub <- cbc_hgb[cbc_hgb$mrn == mrn, ]
  fst_hgb <- sub[sub$taken_time == min(sub$taken_time),]$ord_value
  cbc_fst_hgb[i,] <- c(mrn, fst_hgb)
  remove(sub)
```

# Load the two datasets
```{r}
df1 <- read.csv("All_roots_data08_04_21.csv") # stanford data 
df2 <- read.csv("STS_data_03282014_12312020.csv") # stanford data submit to STS 2014 - 2020
df3 <- read.csv("CompleteAorticRootSu-20142020RootData_DATA_2021-08-18_1812.csv") # stanford root 2014-2020
vartab <- read.csv("v2.9_var.csv")

# df1 and df3 are the same variables, rbind the two, remove duplicated records

df1$redcap_data_access_group <- NULL # remove this variable that does not exist in df3
df1 <- rbind(df1[!df1$mrn %in% df3$mrn,], df3) %>% as.data.frame()
```

# Subset variables and make 
```{r}
conVars <- vartab[vartab$stanford_dataset_varnames == "consistent" & !is.na(vartab$stanford_dataset_varnames),]$variable %>% as.character() 
nonVars <- vartab[vartab$stanford_dataset_varnames == "NA",]$variable %>% as.character() 
harmVars <- vartab[!is.na(vartab$stanford_dataset_varnames) & vartab$stanford_dataset_varnames != "consistent",]$stanford_dataset_varnames %>% as.character()
newVarNames <- vartab[vartab$stanford_dataset_varnames %in% harmVars,]$variable %>% as.character()
```

# Harmonize variables (make stanford data look like the sts data)
```{r}
# diabetes
df1$diabetes_hx[df1$diabetes_hx == 0] <- 2
df1$diabetes_hx[is.na(df1$diabetes_hx)] <- 3

# dislipidemia
df1$hld[df1$hld == 0] <- 2
df1$hld[is.na(df1$hld)] <- 3

# hypertension
df1$htn[df1$htn == 0] <- 2
df1$htn[is.na(df1$htn)] <- 3

# endocarditis 
df1$endocarditis[df1$endocarditis == 0] <- 2
df1$endocarditis[is.na(df1$endocarditis)] <- 3

# chronic lung disease, make sts binary look like stanford data
df2$chrlungd[df2$chrlungd == 1] <- 0
df2$chrlungd[df2$chrlungd !=0 & !is.na(df2$chrlungd)] <- 1
df2$chrlungd[df2$chrlungd == 0] <- 2
df2$chrlungd[is.na(df2$chrlungd)] <- 3
df1$lung_dz_hx[df1$lung_dz_hx == 0] <- 2
df1$lung_dz_hx[is.na(df1$lung_dz_hx)] <- 3

# pvd 
df1$pad[df1$pad == 0] <- 2
df1$pad[is.na(df1$pad)] <- 3

# aortic disease
df1$type_a[df1$type_a == 0] <- 2
df1$type_a[is.na(df1$type_a)] <- 3

# CVD 
df1$hx_cva[df1$hx_cva == 0] <- 2
df1$hx_cva[is.na(df1$hx_cva)] <- 3

# aortic insufficiency
df1$aortic_insufficiency <- df1$aortic_insufficiency - 1
df1$aortic_insufficiency[is.na(df1$aortic_insufficiency)] <- 5

# aortic stenosis
df1$aortic_stenosis[df1$aortic_stenosis == 99] <- NA
df1$aortic_stenosis[df1$aortic_stenosis != 98 & !is.na(df1$aortic_stenosis)] <- 1
df1$aortic_stenosis[df1$aortic_stenosis == 98] <- 2

# genetic connective tissue disease
df1$ctd_hx[df1$ctd_hx == 1] <- 8
df1$ctd_hx[df1$ctd_hx == 2] <- 1
df1$ctd_hx[df1$ctd_hx == 99] <- 7

# prior aortic surgery
df1$aortic_surg_hx <- 2
df1$aortic_surg_hx[df1$aortic_surg_hx___1 == 1 | df1$aortic_surg_hx___2 == 1 | df1$aortic_surg_hx___3 == 1 | df1$aortic_surg_hx___4 == 1] <- 1

# circ arrest
df1$circ_arrest[df1$circ_arrest == 0] <- 2

# long ventilation > 24 hrs
df1$long_vent[df1$long_vent == 0] <- 2

# transient stroke
df1$cnstrokttia[df1$cnstrokttia == 0] <- 2

# new dialysis
df1$hd[df1$hd == 0] <- 2

# decrease date, modify the sts format in this case, from mm/dd/yyyy to yyyy-mm-dd
df2$mtdate <- as.Date(df2$mtdate, "%m/%d/%Y")
df2$dob <- as.Date(df2$dob, "%m/%d/%Y")
df2$dischdt <- as.Date(df2$dischdt, "%m/%d/%Y")


# create 2 holder variables 
df1$mt30stat <- NA
df1$dischdt <- NA

# recode sex
df1$male[df1$male == 0] <- 2 

# make one categorical race variable
df1$race <- NA
df1$race[df1$racecaucasian == 1] <- 1
df1$race[df1$raceblack == 1] <- 2
df1$race[df1$raceasian == 1] <- 3
df1$race[df1$racenativeam == 1] <- 4
df1$race[df1$racnativepacific == 1] <- 5
df1$race[df1$raceother == 1] <- 6

df2$race <- NA
df2$race[df2$racecaucasian == 1] <- 1
df2$race[df2$raceblack == 1] <- 2
df2$race[df2$raceasian == 1] <- 3
df2$race[df2$racenativeam == 1] <- 4
df2$race[df2$racnativepacific == 1] <- 5
df2$race[df2$raceother == 1] <- 6

# generate placeholder variable in df1 for transfusion protocol
df1$transfalg <- NA

df1$creatlst <- NA
df1$postcreat <- NA
```

# Change variable names
```{r}
setnames(df1, old = harmVars, new = newVarNames)
```

# Generate procedure code for the datasets
```{r}
df1$aorta <- 1
df1$cabg <- 0
df1$valve <- 0
df1$other <- 0
df1$afib <- 0
df1$noncard <- 0

df2$aorta <- ifelse(df2$aortproc == 2, 0, 1)
df2$cabg <- ifelse(df2$opcab == 2, 0, 1)
df2$valve <- ifelse(df2$opvalve == 1, 1, 0)
df2$other <- ifelse(df2$opocard == 2, 0, 1)
df2$afib <- ifelse(df2$afibproc == 1, 1, 0)
df2$noncard <- ifelse(df2$oponcard == 1, 1, 0)
```

# Merge the two dataset
```{r}
# Harmonize the rest of the dates
dateVars <- c("dob", "surgdt", "dischdt", "mtdate")
numVars <- c(newVarNames[!newVarNames %in% c(dateVars, "medrecn", "gender")], conVars[!conVars %in% dateVars])

df1[,dateVars] <- lapply(df1[,dateVars], as.Date)
df2[,dateVars] <- lapply(df2[,dateVars], as.Date)

# Generate 30-day mortality variable for the stanford data
df1$mt30stat <- ifelse(lubridate::time_length(difftime(df1$surgdt, df1$dob), "years") <= 30, 2, 1)

# Generate death before discharge
df2$operative_mortality <- ifelse(df2$mtdate < df2$dischdt, 1, 0)

# Generate the final mortality outcome variable
df1$opMort <- ifelse(df1$operative_mortality == 1 | df1$mt30stat == 2, 1, 0)
df2$opMort <- ifelse(df2$operative_mortality == 1 | df2$mt30stat == 2, 1, 0)

addedVars <- c("aorta", "cabg", "valve", "other", "afib", "noncard", "opMort", "race")

# Harmonize the factor variables
df1[,c("postophemoglobin", "postophct")] <- lapply(df1[,c("postophemoglobin", "postophct")], function(x) as.numeric(as.character(x)))
df2[,c("postophemoglobin", "postophct")] <- lapply(df2[,c("postophemoglobin", "postophct")], function(x) as.numeric(as.character(x)))

# Remove the patients from stanford data that are already in sts
df <- rbind(df1[!df1$medrecn %in% df2$medrecn,c(conVars, newVarNames, addedVars)], df2[,c(conVars, newVarNames, addedVars)]) %>% as.data.frame()
```

# Duplicated record choose the most recent operation (There are 175 duplcated record)
```{r}
mrn_dup <- df$medrecn[duplicated(df$medrecn)]

df <- df[order(df$medrecn, df$surgdt, decreasing = T),]
df <- df[!duplicated(df$medrecn),] # will keep the most recent dates
```

# Merge the lowest postoperative hgb
```{r}
cbc_lwst_hgb$medrecn <- cbc_lwst_hgb$mrn
df <- merge(df, cbc_lwst_hgb, by = "medrecn")
```

# Merged added first postop hgb 08/26/2021
```{r}
df <- read.csv("bloodTransfusionMergedSet_08112021.csv", check.names = F)
cbc_fst_hgb$medrecn <- cbc_fst_hgb$mrn
df <- merge(df, cbc_fst_hgb, by = "medrecn")
```

## Compute eGFR
```{r}
df <- read.csv("bloodTransfusionMergedSet_0922021.csv", check.names = F)
df$sex <- ifelse(df$gender == 2, 1, 0) # code women as 1 for easy computation
df$black <- ifelse(df$race == 2, 1, 0) # code black as 1 for easy computation

df_nas <- df[is.na(df$creatlst) | is.na(df$race),]
df_nas$egfr <- NA

df <- df[!df$medrecn %in% df_nas$medrecn,] # remove records with missing values
df$egfr <- 175 * (df$creatlst)^(-1.154) * (df$age)^(-0.203) # male nonblack
df[df$sex==1&df$black==1,]$egfr <- df[df$sex==1&df$black==1,]$egfr*0.742*1.21
df[df$sex==1&df$black==0,]$egfr <- df[df$sex==1&df$black==0,]$egfr*0.742
df[df$sex==0&df$black==1,]$egfr <- df[df$sex==0&df$black==1,]$egfr*1.21

df <- rbind(df, df_nas) %>% as.data.frame()

df$egfrCat <- NA
df$egfrCat[df$egfr>= 90] <- "Stage 1"
df$egfrCat[df$egfr>= 60 & df$egfr <90] <- "Stage 2"
df$egfrCat[df$egfr>= 45 & df$egfr <60] <- "Stage 3A"
df$egfrCat[df$egfr>= 30 & df$egfr <45] <- "Stage 3B"
df$egfrCat[df$egfr>= 15 & df$egfr <30] <- "Stage 4"
df$egfrCat[df$egfr< 15] <- "Stage 5"
df$egfrCat <- as.factor(df$egfrCat)
```

# Save the merged dataset
```{r}
write.csv(df, file = "bloodTransfusionMergedSet_09222021.csv", row.names = F)
```

# Variable request from other instutions
```{r}
vartab <- read.csv("v2.9_var.csv")

# Must haves
mustVars <- vartab[vartab$Needs == "yes",]$variable %>% as.character()
goodToHaveVars <- vartab[vartab$Needs == "maybe",]$variable %>% as.character()

write.csv(mustVars, file = "dataRequestMustHaveVars_08282021.csv", row.names = F)
write.csv(goodToHaveVars, file = "dataRequestgoodToHaveVarsVars_08282021.csv", row.names = F)
```

# 10/06/2021 Add EJ
```{r}
df <- read.csv("bloodTransfusionMergedSet_09222021.csv")
df_new <- read.csv("STS_data_03282014_12312020_10062021.csv")

df <- merge(df, df_new[,c("medrecn", "prelvef")]) %>% as.data.frame()
write.csv(df, file = "bloodTransfusionMergedSet_09222021.csv", row.names = F)
```

# 10/06/2021 Check columbia data
```{r}
library(readxl)
df_c <- read_excel("TRACRequestA9.27.xlsx")
```








